Article Title

### 2013-12-26 plos.tex
completed Methods -- still needs some refining.

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations in neural activity measured in such recordings can reveal important aspects of the functional organization of neural circuits.  However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.  Estimation can be improved by regularization: biasing of the usual unbiased estimate toward a low-dimensional approximation.  The amount of improvement is greatest when the approximation chosen as the target of regularization parsimoniously captures the dominant dependencies in the data.  Therefore, the selection of the most efficient estimator of neural correlation matrices is an empirical question that informs about the types of dominant dependencies governing the system.

@@

\paragraph{Illustration of estimation of a neural correlation matrix}
To illustrate the challenges in estimating the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}. Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero. We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (\figref{01}{A, B}) and covariance selection (\figref{01}{C}). Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales. The close similarity of the two estimates of the correlation matrix (\figref{ 01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients. Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth. Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

@@

We examined the structure of the sparse+latent estimate at individual sites (an example site is depicted in Fig.\;\ref{fig:03}). At these sites the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (\figref{05}{A and D}). However, the corresponding partial correlation matrices differed dramatically (\figref{05}{B and E}). The partial correlation was decomposed into sparse and low-rank components (\figref{ 05}{C}). Although correlations were mostly positive, the sparse partial correlations (or `interactions'), had a much larger fraction of negative values than sample correlations. The sparse component had 82.1\% sparsity (or, conversely, 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (\figref{05}{G}). The low-rank component had rank 17.

@@

We also examined the relationship of differences in orientation preference and physical distances (lateral and by depth) between pairs of cells to the average sample correlations, average regularized partial correlations, and inferred connectivity between them (Fig.\;\ref{fig:07}). The average partial correlations fell more rapidly with difference in preferred orientation (\figref{fig:07}{A and D}) and lateral displacements at equal depths ($\pm 25\;\mu$m) (\figref{07}{B and E}), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (\figref{07}{C and F}).

@@

\paragraph{Big neural data}
Transformational breakthroughs in systems neuroscience are anticipated to follow the ongoing technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far \cite{Alivisatos:2013}. It is far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques. Large datasets aggravate the numerical problems and the model selection caveats described in this paper. The number of parameters in models of population activity grows quadratically.  Models of joint population activity feature large numbers of parameters, that growing quadratically or faster with population size. Without ways to restrict such models, they quickly become ill conditioned (e.g. \cite{Roudi:2009}.  Statistical models of population activity will need to have both correct mathematical forms of interactions and optimal ways of restricting them so that the total number of estimated parameters remains commensurate with the amount of measurements.

@@

\section*{Methods}
% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\paragraph{Data Acquisition}
Fast 3D two-photon imaging of calcium signals \emph{in vivo} was performed as described in \cite{Cotton:2013}

\paragraph{Cross-validation}
To compare the performance of the estimator against each other, we used conventional 10-fold cross validation to measure cross-validation loss (Eq.~\ref{eq:cv-loss}). Briefly, each recording was split into 30 blocks of equal duration.  The 30 blocks were then grouped randomly into 10 datasets with 3 blocks in each.  This procedure ensured sufficient independence between the 10 datasets while still ensuring that each dataset included data from different parts of the recording.   Then, each dataset was used as the testing dataset with the rest of the data used for estimating the covariance matrix.

@@

For plots in \figref{05}{B and E}, \figref{06}{A}, and \figref{07}{A--C}, the regularized partial correlations were obtained by normalizing the precision matrix (Eq.~\ref{eq:partial}). The sparse and low-rank components in \figref{05}{C} are the additive components of these normalized partial correlations.

The sparse partial correlations computed for the display of connectivity in \figref{05}{F, G, H}, was computed by normalizing only the sparse component of the precision matrix still using Eq.~\ref{eq:partial}.

@@

\paragraph{Simulation}

\section*{Acknowledgments}
We Genevera Allen

Eftychios Pnevmatikakis

@@

%\section*{Figure Legends}
\newpage
\section*{Supplementary Figures}
\setcounter{figure}{0}
\begin{figure}[!ht]
\begin{center}
\includegraphics{./figures/Figure-Supp01.pdf}
\end{center}
\caption{{\bf All-to-all performance comparisons of the sample covariance matrix and the four regularized estimators.}
}
\label{supp:01}
\end{figure}
\end{document}

>>>>

\section*{Abstract}
% Please keep the abstract between 250 and 300 words
Ambitious projects aim to record the activity of ever larger and denser subsets of neurons \emph{in vivo}.  Correlations in neural activity measured in such recordings can reveal important aspects of the functional organization of neural circuits.  However, estimating and interpreting large correlation matrices obtained from finite recordings is challenging.  Estimation can be improved by regularization: imposing some kind of structure on the estimate.  The amount of improvement depends on how well the imposed structure accommodates the dominant dependencies in the data.  Therefore, the selection of the most efficient estimator of neural correlation matrices is an empirical question that informs about the types of dominant dependencies governing the system.

@@

\paragraph{Illustration of estimation of a neural correlation matrix}
To illustrate the challenges in estimating the correlation matrix from a finite sample, we consider a regularization scheme based on \emph{covariance selection} (Figure \ref{fig:01}) \cite{Dempster:1972}. Here the estimate is produced by fitting only an optimal subset of the coefficients of the precision matrix while setting the rest to zero. We estimated the covariance matrix of somatic calcium signals of a group of 298 neurons in mouse visual cortex using both the sample covariance matrix (\figref{01}{A, B}) and covariance selection (\figref{01}{C}). Due to the low-pass filtration effect of calcium dye kinetics, correlations in unprocessed calcium signals are higher than typical firing rate correlations on shorter temporal scales. The close similarity of the two estimates of the correlation matrix (\figref{01}{B and C}) belies the utterly different partial correlation structure (\figref{01}{D and E}): The regularized estimate is produced from the same data by setting to zero 31,501 (71.2\%) of the off-diagonal coefficients of the precision matrix and fitting the remaining 12,752 coefficients. Estimation by covariance selection is attractive for several reasons: First, with fewer free parameters it is less susceptible to sampling noise, yielding estimates that are, on average, closer to the truth. Additionally, if we assume that partial correlations do indeed reflect elements of the functional connectivity in the circuit, the second estimate also appears more plausible: in neural circuits, each neuron only interacts with a fraction of the entire population.

@@

We examined the structure of the sparse+latent estimate at individual sites (an example site is depicted in Fig.\;\ref{fig:03}). At these sites the regularized estimate of the correlation matrix appeared very similar to the unregularized sample correlation matrix (\figref{05}{A and D}). However, the corresponding partial correlation matrices differed dramatically (\figref{05}{B and E}). The partial correlation was decomposed into sparse and low-rank components (\figref{05}{C}). Although correlations were mostly positive, the sparse partial correlations (or `interactions'), had a much larger fraction of negative values than sample correlations. The sparse component had 82.1\% sparsity (or, conversely, 17.9\% connectivity), which corresponded to the average node degree (interactions per cell) of 52.5 (\figref{05}{G}). The low-rank component had rank 17.

@@

We also examined the relationship of differences in orientation preference and physical distances (lateral and by depth) between pairs of cells to the average sample correlations, average regularized partial correlations, and inferred connectivity between them (Fig.\;\ref{fig:07}). The average partial correlations fell more rapidly with difference in preferred orientation (\figref{07}{A and D}) and lateral displacements at equal depths ($\pm 25\;\mu$m) (\figref{07}{B and E}), and differences in depth at small ($\pm 25\;\mu$m) lateral displacements (\figref{07}{C and F}).

@@

\paragraph{Big neural data}
Transformational breakthroughs in systems neuroscience are anticipated to follow the ongoing technological advances that will allow simultaneous recordings of the spiking activity of much larger subsets of neurons in functioning neural circuits than have been possible thus far \cite{Alivisatos:2013}. It is far from certain whether massively multineuronal recordings will readily reveal new principles of neural function by application of current statistical techniques. Large datasets aggravate the numerical problems and the model selection caveats described in this paper. The number of parameters in models of population activity grows quadratically.  Models of joint population activity feature large numbers of parameters, that growing quadratically or faster with population size. Without ways to restrict such models, they quickly become ill conditioned (e.g. \cite{Roudi:2009}).  Statistical models of population activity will need to have both correct mathematical forms of interactions and optimal ways of restricting them so that the total number of estimated parameters remains commensurate with the amount of measurements.

@@

\section*{Methods}
% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\paragraph{Surgery and two-photon Imaging}
All procedures were conducted in accordance with the ethical guidelines of the National Institutes of Health and were approved by the Baylor College of Medicine IACUC.  The surgical procedures and data acquisition were performed as descrbed in \cite{Cotton:2013}. Briefly, C57BL/6J mice (aged p40--60) were used. Anesthesia was initiated with isoflurane (3%) and the mixture of fentanyl (0.05 mg/kg), midazolam (5 mg/kg), and medetomidine (0.5 mg/kg), with boosts of half the initial dose every 3 hours.  A craniotomy was performed over the right primary visual area.  Membrane-permeant calcium indicator Oregon Green 488 BAPTA-1 AM (OGB-1, Invitrogen) was loaded by bolus injection.  The craniotomy was sealed using a glass coverslip secured with dental cement.

Calcium imaging began 1 hour after dye injection.  All imaging was performed using the 3D-RAMP two-photon microscope described in \cite{Cotton:2013}. First, A 3D stack was acquired and cells manually segmented.  To collect calcium signals, the system repeated hopped between the selected neurons. 
\paragraph{Visual stimulus}
Full-field drifting gratings with 90\% contrast, luminance of 10 cd/m$^2$, spatial frequency of 0.08 cycles/degree, and temporal frequency of 2 cycles/s. Two sets of stimuli were presented for each imaging site: the first to map directional tuning and the second to estimate noise correlations. Directional tuning was map using a pseudo-random sequence of drifting gratings at sixteen equally spaced directional of motion changing at 2 Hz for 3 min without blanks. The data for covariance estimation were collected during presentations of full-field drifting gratings with the same parameters as those used in directional tuning except only two directions (in 9 dataset) or five directions (in 22 datasets) were used and the presentations lasted 1 second and, separated by 1-second blanks.  Each stimulus condition was presented at least 180 times. 
\paragraph{Data processing}
All data were processed in MATLAB using the DataJoint data processing chain toolbox ({\tt https://github.com/datajoint/}) developed in our lab.

The collected fluorescent traces were deconvolved to reconstruct the firing rates for each neuron. First, the first principal component was subtracted from the traces, which reduced common mode noise related to small movement artifacts and cardiovascular artifacts. The resulting traces were low-pass filtered below 0.1 Hz and downsampled to 20 Hz. Firing rates were estimated using a fast non-negative deconvolution algorithm \cite{Vogelstein:2010}.

Orientation tuning was computed by fitting the mean firing rates in response to gratings of directions $\phi$ with two-peaked von Mises tuning functions of the form $f(\phi)=a + b\exp\left[(\cos(\phi-\theta)-1)/w\right] + c\exp\left[(-\cos(\phi-\theta)-1)/w\right]$ where $b\ge c$ are amplitudes of the two respective peaks, $w$ is the tuning width, and  $\theta$ is the preferred direction. The significance of the fit was determined by the permutation test: the labels of the direction were randomly permuted 10,000 times.  The $p$-value of the fit was computed as the fraction of permuted datasets for which the $R^2$ value of the tuning function fit exceeded that of the real data.  Cells were considered tuned for $p$-values not exceeding $0.05$.

For covariance estimation, the analysis was limited to the period with 2 or 5 stimulus conditions and lasted between 12 and 20 mins.  Only traces whose average deconvolved signal was greater than 1\% of the median of the recorded population were included in the analysis.  
\paragraph{Cross-validation}
To compare the performance of the estimator against each other, we used conventional 10-fold cross validation to measure cross-validation loss (Eq.~\ref{eq:cv-loss}). Briefly, each recording was split into 30 blocks of equal duration.  The 30 blocks were then grouped randomly into 10 datasets with 3 blocks in each.  This procedure ensured sufficient independence between the 10 datasets while still ensuring that each dataset included data from different parts of the recording.   Then, each dataset was used as the testing dataset with the rest of the data used for estimating the covariance matrix.

@@

The partial correlation matrix computed from the overall estimate, computed from $C_{\sf sparse+lowrank}$ according to Eq.~\ref{eq:partial}, includes the effects of interactions between the visible and latent units.  This estimate of the partial correlations was used in analyses where sparsity was not required (\figref{05}{B and E} and \figref{06}{A}, and \figref{07}{A--C}).  For analyses that required a sparse network (\figref{05}{F, G, H} and \figref{06}{B, D} and \figref{07}{D--F}) was computed using the extended precision matrix $S^\ast$:
\begin{equation}
P_{\sf sparse} = (S\circ I)^{-\frac 1 2} S  (S\circ I)^{-\frac 1 2}
\end{equation}

@@

\paragraph{Simulation}
For simulation, ground truth covariance matrices were produced by taking 150 independent samples from an artificial population of 50 independent, identically normally distributed units. The covariance matrices were then subjected to the respective regularizations to produce the ground truth matrices for the simulation studies (\figref{02}{\,row 2}. Samples were then generated by multivariate normal distributions with the respective true covariance matrices and estimated by each of the estimators.

\section*{Acknowledgments}
We thank Genevera Allen for a helpful discussion.  We thank Eftychios Pnevmatikakis for helpful suggestions and feedback.

@@

%\section*{Figure Legends}
\newpage
\section*{Supporting Information}
\setcounter{figure}{0}

\begin{figure}[!ht]
\begin{center}
\includegraphics{./figures/Figure-Supp01.pdf}
\end{center}
\caption{{\bf All-to-all performance comparisons of the sample covariance matrix and the four regularized estimators.}
}
\label{supp:01}
\end{figure}

\end{document}
